%-------------------------------------------------------------------
\chapter{Related Works}
\label{ch:sota}
%-------------------------------------------------------------------

\epigraph{\enquote{The perception of depth in a monocular picture is based completely upon the assumptions of the observer.}}{\emph{Lawrence Roberts}}

This chapter provides an overview of the methods that researchers developed for tackling the single image depth estimation (SIDE) problem.
The developed approaches are so many that it is not feasible to faithfully report all of them.
Hence, a selection has been made.

The chapter is divided into the following sections:
\begin{enumerate}
    \item{Datasets: an overview on datasets used for SIDE.}
    \item{Metrics and Evaluation: which metrics and evaluation procedures are used in literature.}
    \item{Regression Methods: the most straightforward approach to the problem.}
    \item{Bin Methods: neural networks are known to be limited in regression tasks, a classification based approach is explored in this section.}
    \item{Miscellaneous: important works that do not fall under the same category.}
    \item{Self Supervised Methods: supervised learning needs labeled data and labeled data is expensive, this section reviews works that use unlabeled data for training SIDE models.}
    \item{Classics: methods based on handcrafted pipelines.}
    \item{Deep Learning and Interpretability: a review of the concept of "interpretability" in the deep learning scenario.}
    \item{Single Image Depth Estimation and Interpretability: overview of relevant works about interpretability in single image depth estimation.}
\end{enumerate}

\input{chapter_sota/chap2.inc}

\input{chapter_sota/SL.inc}

\input{chapter_sota/SSL.inc}

\input{chapter_sota/SOTA.inc}

\input{chapter_sota/classics.inc}